Enzo+Moray: Radiation Hydrodynamics Adaptive Mesh Refinement Simulations with Adaptive Ray Tracing

### 2010-08-14 ms.tex
Added missing references.
Adding Test 12 figure of kph and description of the radiation pulses.

The absorption of radiation transfers momentum from photons to the
absorbing medium, i.e. radiation pressure.  This is easily computed by
considering the momentum
\begin{equation}
  d\mathbf{p}_\gamma = \frac{dP \; E_{\rm ph}}{c\;dt} \; \hat{r}
\end{equation}
of the absorbed radiation from the incoming ray, where $\hat{r}$ is
the normal direction of the ray and $dt$ is the radiative transfer
timestep.  The resulting acceleration of the gas due to radiation
pressure is
\begin{equation}
  d\mathbf{a} = d\mathbf{p}_\gamma / (\rho \; V_{\rm cell}),
\end{equation}
where $\rho$ is the gas density inside the cell.  This acceleration is
then added to the other forces, e.g. gravity and gas pressure, in the
calculation.

@@

Our implementation retains the time derivative of the radiative
transfer equation (Eq. \ref{eqn:rteqn}) if we choose a constant ray
tracing timestep, which saves the photon packages between timesteps if
the $c\;dt_{\rm P} < L_{\rm box}$.  This effect only becomes apparent
when the variation timescale of the point source is smaller than the
light crossing time of the simulation.  Furthermore, the timestep
should resolve the variation timescale by a few times.  To test this,
we can use an exponentially varying source with some duty cycle.  In a
functional form, this can be described as
%
\begin{equation}
  \label{eqn:test13}
  L(t) = L_{\rm max} \times \exp(t_f/t_0 - 1),
\end{equation}
%
where $t_f = \mathrm{mod}(t,t_0)$ and $t_0$ is the duty cycle.  To
illustrate the variations in the photo-ionization rates due to a
varying source, we remove any dependence on the medium by considering
an optically-thin uniform density.  We take $L_{\rm box} = 1$ Mpc,
which has a light crossing time of 3.3 Myr.  Thus we consider a source
at the origin with $L_{\rm max} = 10^{50}$ ph s$^{-1}$ and $t_0 = 0.5$
Myr.  We use a radiative transfer timestep of 50 kyr to resolve the
duty cycle by 10 timesteps.

>>>>

The absorption of radiation transfers momentum from photons to the
absorbing medium, i.e. radiation pressure.  This is easily computed by
considering the momentum
\begin{equation}
  d\mathbf{p}_\gamma = \frac{dP \; E_{\rm ph}}{c} \; \hat{r}
\end{equation}
of the absorbed radiation from the incoming ray, where $\hat{r}$ is
the normal direction of the ray.  We do not include radiation pressure
on dust, and thus our treatment underestimates its effects in
applicable environments.  The resulting acceleration of the gas due to
radiation pressure is
\begin{equation}
  d\mathbf{a} = \frac{d\mathbf{p}_\gamma} {dt\; \rho \; V_{\rm cell}},
\end{equation}
where $\rho$ is the gas density inside the cell and $dt$ is the
radiative transfer timestep.  This acceleration is then added to the
other forces, e.g. gravity and gas pressure, in the calculation.

@@

\begin{figure}[t]
  \epsscale{0.6}
  \plotone{fig/test12_kph}
  \caption{\label{fig:test12_kph} Test 12 (Time variations of the
    source luminosity).  Slice of the photo-ionization rate
    \kph~through the origin.  The source has a duty cycle of 0.5 Myr,
    and the box has a light crossing time of 3.3 Myr.  The shells of
    high \kph~originate from radiation that was emitted when the
    source was at its peak luminosity, illustrating the
    time-dependence of the radiative transfer equation.}
  \epsscale{1}
\end{figure}

Our implementation retains the time derivative of the radiative
transfer equation (Eq. \ref{eqn:rteqn}) if we choose a constant ray
tracing timestep, which saves the photon packages between timesteps if
the $c\;dt_{\rm P} < L_{\rm box}$.  This effect only becomes apparent
when the variation timescale of the point source is smaller than the
light crossing time of the simulation.  Furthermore, the timestep
should resolve the variation timescale by a few times.  To test this,
we can use an exponentially varying source with some duty cycle.  In a
functional form, this can be described as
%
\begin{equation}
  \label{eqn:test13}
  L(t) = L_{\rm max} \times \exp[A(t_f/t_0 - 1)]
\end{equation}
%
where $t_f = 2 \times |t - t_0 \times \mathrm{round}(t/t_0)|$, $t_0$
is the duty cycle, and $A=4$ controls the width of the radiation
pulse.  To illustrate the variations in the photo-ionization rates,
shown in Figure \ref{fig:test12_kph}, due to a varying source, we
remove any dependence on the medium by considering an optically-thin
uniform density $\rho 10^{-4}~\cubecm$.  We take $L_{\rm box} = 1$
Mpc, which has a light crossing time of 3.3 Myr.  A source is placed
at the origin with $L_{\rm max} = 10^{55}$ ph s$^{-1}$ and $t_0 = 0.5$
Myr.  We use a radiative transfer timestep of 50 kyr to resolve the
duty cycle by 10 timesteps.
